### 2013-06-10 parts/asp.tex
new section for ASP

\section{Answer Set Programming implementation concepts}

\rewrite{Check and correctly glue this section}

\newcommand{\ti}[1]{\texttt{\textit{#1}}}
\newcommand{\aspil}[1]{\texttt{#1}}
\newcommand{\asp}[1]{\begin{itemize} \item[] \aspil{#1} \end{itemize}}

%\newcommand{\atom}{\mathbf}
\newcommand{\atom}[1]{#1}
%\newcommand{\predicate}{\mathbf}
\newcommand{\predicate}[1]{#1}
\newcommand{\la}{\leftarrow}
\newcommand{\var}[1]{#1}
\newcommand{\nota}{\neg}

\newcommand{\paramlabel}{\predicate{param\_label}}
\newcommand{\paramres}{\predicate{param\_resource}}
\newcommand{\component}{\predicate{component}}
\newcommand{\componentlevels}{\predicate{component\_levels}}
\newcommand{\param}{\predicate{param}}
\newcommand{\inferedparam}{\predicate{infered\_param}}
\newcommand{\lessactive}{\predicate{less\_active}}
\newcommand{\paraminf}{\predicate{param\_inf}}

Answer Set Programming (ASP) is a logic programming paradigm \cite{Baral03},
which has been chosen to address the enumeration of all admissible parametrizations.
The motivations are following:
\begin{itemize}
  \item ASP efficiently tackles the inherent complexity of the models used, thus allowing a fast execution of the formal tools defined in this paper,
  \item ASP is convenient to enumerate a large set of possible answers,
  \item and it allows to easily constrain the answers according to some properties.
\end{itemize}
We now synthesize some key points to better make the reader understand our ASP implementation with the enumeration example.

\subsection{Simple rules}\label{sssec:simple_rules}
ASP is based on a set of rules of of the form:
\begin{align*}
  \underbrace{{\ }\atom{H}_{\ }}_{head} \la \underbrace{\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m}_{body}.
\end{align*}
where the $body$ is a series of atoms ($\atom{A}_i$) and negations of atoms ($\nota \atom{B}_i$).
In the case of \emph{simple rules} (as opposed to the \emph{cardinality rules} of \pref{sssec:cardinality_rules}), the $head$ is also an atom ($\atom{H}$).
Such a rule states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
and all atoms $\atom{B}_1, \atom{B}_2, \dots, \atom{B}_m$ are not true (negation by failure), then $\atom{H}$ has to be true.
Solving an ASP program means finding an \emph{answer set}, that is a minimal set of true atoms that respect all the rules.
Several answer sets can be solution to the same ASP program and the solver can be directed to enumerate them all.

An atom is composed of a predicate and a series of arguments (possibly empty).
For example, the following atom:
\begin{align*}
  \predicate{p}(x_1, x_2, \dots, x_r)
\end{align*}
is composed of the predicate $\predicate{p}$ and $r$ arguments: $x_1, x_2, \dots, x_r$.
Each argument can be either an value or a variable,
a value being a representation of a piece of data (component name, expression level, …)
and a variable can is used to represent any existing value which respects the rules.
In this paper, a variable is always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, …)
while values are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …).

A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
For instance, the information describing the studied model (the original PH model and the inferred IG and parameters) are expressed in ASP using facts.
In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and that $l_a = m$.

\begin{example}
Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
\begin{align*}
  &\component(a, 2). \\
  &\component(b, 1). \\
  &\component(c, 1).
\end{align*}
These three atoms are built with the predicate $\component$.
Furthermore, $a$, $b$, $c$, $1$ and $2$ are values.
\end{example}

To describe the sets of all expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),
one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
Variables here come in handy to enumerate each possible value $k$ for each component $a$:
during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible values it could represent.
The following rule, for example, contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$) and enumerates the set of possible expression levels of each component in the system:
\begin{align*}
  \componentlevels(A, K) \la \component(\var{A}, \var{M}), 0 \leq K \leq M.
\end{align*}
where the notation “$\leq$” stands for a shortcut in ASP which has the same meaning as the mathematical operator.

\begin{example}
Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to every answer set:
\begin{align*}
  &\{&\componentlevels(b, 0)
  &&\componentlevels(b, 1) \\
  &&\componentlevels(c, 0)
  &&\componentlevels(c, 1) \\
  &&\componentlevels(a, 0)
  &&\componentlevels(a, 1) \\
  &&\componentlevels(a, 2) &&&&\}
\end{align*}
\end{example}

\subsection{Cardinality rules}
\label{sssec:cardinality_rules}
As an extension of simple rules, \emph{cardinality rules} turn out to be convenient to enumerate a set of answer sets.
The head of a cardinality rule specifies a set of atoms $H$ and two integers $min$ and $max$, and is denoted:
\begin{align*}
  min\ \{\ H\ \}\ max \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
\end{align*}
Given such a rule, as many answer sets as possible are created, so that each answer $S$ set verifies:
\begin{align*}
  min \leq |S \cap H| \leq max
\end{align*}
and every atom $\atom{H}_i \in S \cap H$ respects the simple rule:
\begin{align*}
  \atom{H}_i \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
\end{align*}
In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
which is a shorthand for “the set of atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
\rewrite{Reformulate?}

Cardinality rules turn out to be convenient to enumerate all possible parametrizations by creating multiple answer sets.
For functional purposes, a unique label is assigned to every possible set of resources of a given component.
Thus, we denote $\omega_p$ the set of resources of a given component $a$ labeled by $p$,
and naturally, $K_{a,\omega_p}$ is the related parameter.
%and in the following we note $K_{a,\omega_p}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
We note that labeling the sets of resources of a component is obviously equivalent to labeling its parameters.
Then, suppose that:
\begin{itemize}
  \item $\paramlabel(a, p)$ states that $p$ is a valid label for a set of resources of component $a$ (and therefore $K_{a,\omega_p}$ is a valid parameter);
  \item $\param(a, p, i)$ states that: $i \in K_{a, \omega_p}$;
  \item $\inferedparam(a, p)$ states that the parameter inference of $K_{a, \omega_p}$ was conclusive (\pref{pps:param_K}).
\end{itemize}
It is thereby possible to enumerate the possible values of all parameters for which \pref{pps:param_K} was not conclusive, with the following cardinality rule:
\begin{align*}
  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) : \componentlevels(\var{A}, \var{I})\ \}\ \infty\ \la \\
  & \qquad\qquad\qquad \paramlabel(\var{A}, \var{P}), \nota \inferedparam(\var{A}, \var{P}).
\end{align*}
Indeed, this rule applies to any possible parameter $\var{P}$ of any component $\var{A}$ ($\paramlabel$) whose value is still unknown ($\nota \inferedparam$),
and states that any expression level $\var{I}$ of this component ($\componentlevels$) can belong to the value of the parameter ($\param$).
Furthermore, the lower bound is $1$, which forces each enumerated parameter to contain at least one value,
but no upper bound is specified ($\infty$) for the size of each parameter (which is bounded by the number of possible expression levels of each gene).
%each parameter contains at least $1$ value, as stated by the lower bound, and has no upper bound ($\infty$) but the number of expression levels of its component.
In other worlds, this cardinality rule creates as many answer sets as there are \emph{candidate} parametrizations
so that if $K_{a, \omega_p}$ could not be inferred by \pref{pps:param_K}, then
$K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$
(thus completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K} or the fact that parameters have to be intervals).

%so that
%inferred parameters with \pref{pps:param_K} keep their value and
%parameters that could not be inferred take any possible value amongst the expression levels of the related component: $K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$.

\subsection{Constraints}\label{sssec:constraints}
%Finally, constraints can be used to filter answer sets that model parametrizations which are impossible or do not respect the assumptions of \pref{ssec:admissible-K}.
Finally, a constraint is a rule with no $head$ part:
\begin{align*}
  \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
\end{align*}
A constraint is satisfied only if its $body$ is not satisfied,
which thus allows to invalidate answer sets containing some unwanted combinations of atoms.
In the scope of parameters enumeration, for example, constraints are especially useful to filter parametrizations
that contain non-interval parameters or that do not respect the assumptions of \pref{ssec:admissible-K}.
Indeed, suppose that:
\begin{itemize}
  \item $\lessactive(a, p, q)$ states that $\omega_p$ is a set of resources of $a$ with (loosely) less activators and more inhibitors than $\omega_q$;
  \item $\paraminf(a, p, q)$ states that: $K_{a,\omega_p} \leqsegm K_{a,\omega_q}$.
\end{itemize}
Then, the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following constraint:
\begin{align*}
  \la \lessactive(\var{A}, \var{P}, \var{Q}), \nota \paraminf(\var{A}, \var{P}, \var{Q}).
\end{align*}
Indeed, this constraint removes all parametrization results where parameters $K_{\var{A},\omega_\var{P}}$ and $K_{\var{Z},\omega_\var{Q}}$ exist
such that $\var{A}$ is less activated by the set of resources $\omega_\var{P}$ than $\omega_\var{Q}$,
but $K_{\var{A},\omega_\var{Q}} \ltsegm K_{\var{A},\omega_\var{P}}$,
thus violating the monotonicity assumption.
Of course, other assumptions can be formulated in the same way.

\begin{example}
In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{b\}}$ and $K_{a,\{c\}}$ could not be inferred by \pref{pps:param_K}.
The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values
(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
\begin{align*}
  (K_{a,\{b\}} ; K_{a,\{c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
\end{align*}
We note that $(\segm{1}{1} ; \segm{1}{1})$ belongs to the enumerated set of admissible parametrizations.
Therefore, this enumeration allows, from the model in \pref{fig:runningPH-1}, to find the behavior of the model completed with a cooperative sort described in \pref{fig:runningPH-2}.
\end{example}

\begin{comment}
---------------

In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
\begin{align*}
  &\paramlabel(a,1).
  &\paramlabel(b,1).\\
  &\paramlabel(a,2).
  &\paramlabel(b,2).\\
  &\paramlabel(a,3).\\
  &\paramlabel(a,4).
  &\paramlabel(c,1).
\end{align*}

A set (such as sets of resources $\omega$) can be defined in ASP with a series of atoms. \rewrite{Find a more precise sentence.}
For example, for a given parameter $K^p_{a,\omega}$, expressing $x \in \omega$ is done by the atom: $\paramres(a, p, x)$.
Then, the sets of resources of \pref{ex:infer-param-runningPH-1} is defined by:
$$\begin{array}{>{\hspace{.8em}}l|>{\phantom{\raisebox{.6em}{xx}}}l}
  \hspace{-.8em}\text{Parameter} & \text{Corresponding ASP fact} \\
\hline
  K^0_{a,\emptyset} & \\
  K^1_{a,\{b\}} & \paramres(a,1,b). \\
  K^2_{a,\{c\}} & \paramres(a,2,c). \\
  K^3_{a,\{b,c\}} & \paramres(a,3,b). \quad \paramres(a,3,c). \phantom{\raisebox{-.8em}{xx}} \\ \hline
  K^0_{b,\emptyset} & \\
  K^1_{b,\{a\}} & \paramres(b,1,a). \phantom{\raisebox{-1em}{xx}} \\ \hline
  K^0_{c,\emptyset} &
\end{array}$$
The absence of an atom $\paramres(a,1,c)$, for example, states that $c$ is not a resource for the parameter $K^1_{a,\{b\}}$ (indeed, $c \notin \{b\}$).

\todo{Continue refactoring.}
\rewrite{Explain variables}

Rules allow more detailed declarations than facts as they have a body (right-hand part below) containing constraints and allowing to use variables, while facts only have a head (left-hand part).
For instance, in order to define the set of expression levels of a component, we declare:
$$\atom{component\_levels}(\var{X}, 0..\var{M}) \la \atom{component}(\var{X}, \var{M}).$$
where the atom $\atom{component}(X, M)$ stands for the existence of a component $X$ with a maximum level $M$.
Considering this declaration, any possible answer for the atom $\atom{component\_levels}$ will be found by binding all possible values of its two terms with all existing $\atom{component}$ facts:
the existence of an answer $\atom{component\_levels}(a, k)$ will depend on the existence of a term $a$, which is bound with $\var{X}$, and an integer $k$, constrained by: $0 \leq k \leq \var{M}$.

Cardinalities are convenient to enumerate all possible parametrizations by creating multiple answer sets.
A cardinality (denoted hereafter with curly brackets) gives any number of possible answers for some atoms between a lower and upper bounds.
For example,
\begin{align*}
  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
  & \qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
\end{align*}
where $\atom{param}(\var{X}, \var{P}, \var{I})$ stands for: $\var{I} \in K^\var{P}_{\var{X},\omega}$,
means that any parameter of component $\var{X}$ and label $\var{P}$ must contain at least one level value ($\var{I}$) in the possible expression levels of $\var{X}$.
Indeed, the lower bound is $1$, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
The body (right-hand side) of the rule also checks for the existence of a parameter of $\var{X}$ with label $\var{P}$,
and constrains that the parametrization inference was not conclusive for the considered parameter (“$\nota$” stands for negation by failure: $\nota \atom{L}$ becomes true if $\atom{L}$ is not true).
Such a constraint gives multiple results as any set of $\atom{param}$ atoms satisfying the cardinality will lead to a new global set of answers.
In this way, we enumerate all possible parametrizations which respects the results of parameters inference,
but completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}.

We rely on integrity constraints to filter only admissible parametrizations.
An integrity constraint is a rule with no head, that makes an answer set unsatisfiable if its body turns out to be true.
Hence, if we suppose that:
\begin{itemize}
  \item the $\atom{less\_active}(a, p, q)$ atom means that $K^p_{a,\omega}$ stands for a configuration with less activating regulators than $K^q_{a,\omega'}$, % (\ie $A \subset A'$),
  \item the $\atom{param\_inf}(a, p, q)$ atom means: $K^p_{a,\omega} \leqsegm K^q_{a,\omega'}$,
\end{itemize}
then the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following integrity constraint:
$$\la \atom{less\_active}(\var{X}, \var{P}, \var{Q}), \nota \atom{param\_inf}(\var{X}, \var{P}, \var{Q}).$$
which removes all parametrization results where parameters $K^\var{P}_{\var{X},\omega}$ and $K^\var{Q}_{\var{X},\omega'}$ exist such that $\var{X}$ is less activated by $\omega$ than $\omega'$ %$A \subset A'$
but $K^\var{Q}_{\var{X},\omega'} \ltsegm K^\var{P}_{\var{X},\omega}$,
thus violating the monotonicity assumption.
Of course, other assumptions can be formulated in the same way.
\end{comment}

This subsection succinctly described how ASP programs come in handy to represent a model and solve all steps of Thomas' modeling inference.
It finds a particularly interesting application in the enumeration of parameters: all possible parametrizations are generated in separate answer sets, and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,
thus reducing the number of interesting parametrizations to be considered in the end.