### 2013-07-05 main.tex
avancées sur l'intro et l'abstract

\documentclass[preprint]{elsarticle}

\usepackage{hyperref}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}  % Maths
\usepackage{amsfonts} % Maths
\usepackage{amssymb}  % Maths
\usepackage{stmaryrd} % Maths (crochets doubles)

\usepackage{url}     % Mise en forme + liens pour URLs
\usepackage{array}   % Tableaux évolués

\usepackage{comment}

\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{pro}{Property~\ref{#1}}
\newrefformat{pps}{Proposition~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{sssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\newrefformat{ex}{Example~\ref{#1}}
\def\pref{\prettyref}

\newdefinition{definition}{Definition}
\newdefinition{property}{Property}
\newdefinition{proposition}{Proposition}
\newdefinition{example}{Example}

\usepackage{tikz}
\newdimen\pgfex
\newdimen\pgfem
\usetikzlibrary{arrows,shapes,shadows,scopes}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}

\input{macros/macros}
\input{macros/macros-ph}
\input{macros/macros-grn}

% Commandes À FAIRE
%\usepackage{color} % Couleurs du texte
\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
\def\updated#1{{\color{magenta}#1}}
%\newcommand{\new}[1]{#1}

% Id est
\newcommand{\ie}{i.e. }

% Césures
\hyphenation{pa-ra-me-tri-za-tion}
\hyphenation{pa-ra-me-tri-za-tions}

\begin{document}

\begin{frontmatter}

\title{Constructing Biological Regulatory Networks from Process Hitting models}
%\\\rewriteil{Or possibly: “Complementarity between PH, IG and Thomas modeling”}

\author[irccyn,nii]{Maxime Folschette}
\ead{Maxime.Folschette@irccyn.ec-nantes.fr}
\author[eth]{Lo\"ic Paulev\'e}
\author[nii]{Katsumi Inoue}
\author[irccyn,nii]{Morgan Magnin}
\author[irccyn]{Olivier Roux}

\address[irccyn]{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e - B.P. 92101 - 44321 Nantes Cedex 3, France.}
\address[nii]{National Institute of Informatics,\\
2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
\address[eth]{BISON group, Automatic Control Laboratory, ETH Zürich\\
Physikstrasse 3, 8092 Zurich, Switzerland.}

\begin{abstract}
\parindent 0.5cm
The Process Hitting (PH) is a recently introduced framework to model concurrent processes.
Its major originality lies in a specific restriction on the causality of actions, which
makes the formal analysis of very large systems tractable.
PH is suitable to model Biological Regulatory Networks (BRNs) with complete or partial
knowledge of cooperations between regulators by defining the most permissive dynamics
with respect to these constraints.

On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.

This paper establishes formal links between these two formalisms, permitting to appreciate their
complementary features.
First, the inference of the Interaction Graph (IG) from a PH models allows a compact representation of
the regulations that are effective for the dynamics.
From this IG, well-known results can be applied to derive global dynamical properties for the
modelled BRN.
Second, the inference of all Thomas models that are compatible with a given PH is tackled.
As the PH allows partial specification of cooperation between components, this inference
emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
cooperations, where Thomas' approach fails because of the combinatorics of parameters.

The inference of corresponding Thomas' models is provided using Answer Set Programming,
which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
\todo{application}
\end{abstract}

\end{frontmatter}

\input{parts/intro}

\section{Frameworks}\label{sec:frameworks}
\input{parts/ph}
\input{parts/grn}

\input{parts/infer-ig}
\input{parts/infer-param}
\input{parts/asp}
\input{parts/example}
\input{parts/discussion}

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

%\input{parts/contributions}

\end{document}