### 2013-07-05 parts/example.tex

\section{Examples}\label{sec:examples}

\rewrite{Reformulate / glue.}

\subsection{Implementation}

The inference method described in this paper has been implemented as part of
\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
tools.
Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
at \url{http://potassco.sourceforge.net}}.
The IG and parameters inference can be performed using the command
\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
The (possibly partial) inferred parametrization will be returned on the standard output.
The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
parameter to the command.

\todo{Check this paragraph}
Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
our method infers the IG and parametrization given in \pref{fig:runningBRN}.
Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
undefined (because of the lack of cooperativity between $b$ and $c$).
In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
parameters with 3 potential values could not be inferred.
If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
admissible parametrizations.

\subsection{Biological application}

The system studied is extracted from a model of ERBB receptor-regulated G1/S transition detailed in~\cite{Sahin09}.
This model represents a chain of reactions that lead to the activation of protein pRB which is responsible for regulating the cell division,
and is therefore essential to prevent cancer development.
Its IG is composed of 20 components and 52 edges.

Three models are created from the original IG, with different levels of precision regarding the cooperation between components.
Model (1) represents a translation of the raw IG into Process Hitting, that is,
without any knowledge of the Boolean rules (and therefore the cooperations) of the components.
Model (2) implements some of the rules based on the results of several knockdown experiments.
Given the results of the experiments, the introduced cooperative sorts to model the most important cooperations,
which were emphasized by the impossibility to reach the production of pRB without the knocked protein.
Model (3) is the totally refined model with all cooperations implemented given the Boolean functions for all components.
Therefore, those three models can be considered as successive refinements of the original and most general one.
The results of the IG and parameters inferences are detailed in \pref{tb:egfr20}
and discussed in the following alongside with details about their construction.

\begin{table}[ht]
~\hfill%\begin{center}
  \begin{tabular}{r|l|l|m{2cm}|m{2.5cm}|m{1.5cm}}
    \textbf{Model} & \textbf{$|E|$} & \textbf{$|K|$} & \textbf{Inferred parameters} & \textbf{Possible\newline models} & \textbf{Fixed points}
  \\\hline\hline
    (1) & $52$ & $196$ & $20$ & $2^{176}\simeq 9.6\cdot10^{52}$ & $0$   % v1_0.ph
  \\\hline
    (2) & $51$ & $192$ & $98$ & $2^{94}\simeq 2.0\cdot10^{28}$ & $0$    % v2_1.ph
  \\\hline
    (3) & $51$ & $192$ & $192$ & $1$ & $3$                              % ori.ph
  \\\hline
  \end{tabular}
\hfill~%\end{center}
  \caption{Results of the IG and parameters inference on three models
  derived the ERBB receptor-regulated G1/S transition model of~\cite{Sahin09}
  with different precisions in the definition of the cooperations.
  Model (1) contains no cooperations between the components.
  Several cooperations were included in model (2) under the form of 14 cooperative sorts.
  Model (3) models all the Boolean functions between components using 22 cooperative sorts.}
  \label{tb:egfr20}
\end{table}

Model (1) encompasses only sole interactions between components, that is,
independent activations or inhibitions of a component on another given the regulations specified in the original IG.
Therefore, the IG inferred from model (1) is exactly the IG ised to create the model, and the inferred parameters
are the parameters for the extreme cases (all activators present and all inhibitors absent).
This first model therefore abstracts a large number of Thomas models as a lot of parameters are left undecided.

In order to build model (2), 14 cooperative sorts were added in order to model the Boolean functions of several components
(consisting of AND and OR operands).
To do so, the following components were noticed due to their importance in the chain of reactions:
CDK4, CDK6, CycD1, ER$\alpha$ and c-MYC.
Indeed, knockdown experiments have been conducted in~\cite{Sahin09}
and the results showed that knocking down these components lead to an important decrease in the production of pRB.
We therefore concluded that these components were involved in other components' Boolean functions
in a way that their knockdown was sufficient to prevent its production (which is typical of AND operands).
In order to reproduce such requirements, the boolean functions of their successors,
that is CDK4, CDK6, prB, p21, p27, IGF1R, MYC, CycD1 and CycE1,
were modeled as cooperative sorts, if needed.
We note that the added cooperations allowed to infer about half the parameters.
However, the number of possible Thomas models that can be inferred from this PH is still significant
as the remaining unknown parameters leave a big number of possibilities.
Furthermore, the inferred IG contains one edge less than the original IG. This is due to the fact that
one of the Boolean function could in fact be simplified in a way that one of the components did not appear anymore.
No edge have therefore been inferred in this case.

Finally, model (3) was build using all the Boolean functions provided in~\cite{Sahin09}.
These functions take the form of 22 cooperative sorts into the model in order to match the desired behavior of the system.
As all cooperations are fully defined in this model, all the parameters are inferred and only one Thomas model can be derived.
We note also that this model is the only one containing at least one fixed point.

\subsection{Computation times and results}

The current implementation can successfully handle large PH models of BRNs found in the literature
such as the ERBB receptor-regulated G1/S transition model from~\cite{Sahin09} presented in the last subsection,
and a T cell receptor model from~\cite{Klamt06} which contains 40 components and 14 cooperative
sorts\footnote{All models mentioned in this section are available as examples distributed with \textsc{Pint}.}.
For each model, IG and parameters inferences are performed together in less than a second
on a standard desktop computer.

Bigger models related to the aforementioned systems were also tested with our implementation:
an other T cell receptor model described in~\cite{SaezRodriguez2007} containing 94 components
and an EGFR/ErbB signaling model described in~\cite{Samaga2009} which containing 104 components.
\todo{Mention how these models were created?}

The composition of all models and the results of the inferences are summarized in \pref{tb:computation}.
We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
Therefore, these models would therefore be more efficiently studied as PH than as BRNs.
finally, we note that the complexity of the method is exponential in the number of regulators of one
component and linear in the number of components.

\begin{table}
\begin{center}
  \begin{tabular}{r|c|c|c|c}
    \textbf{Model} & \textbf{Components} & \textbf{\# CS} & \textbf{$\Delta t$ IG} & \textbf{$\Delta t$ K}
\\\hline\hline
    ERBB-G1/S & $20$ & $22$ & <1s & <1s
\\\hline
    T cell receptor & $40$ & $14$ & <1s & <1s
\\\hline
    T cell receptor & $94$ & $39$ & 10s & 
\\\hline
    EGFR/ErbB signaling & $104$ & $89$ & 3min 20s & 
  \end{tabular}
\end{center}
  \caption{Other models \rewrite{Caption}}
  \label{tb:computation}
\end{table}

\begin{comment}
\subsection{\warnul{Old version of this section}}

The inference method described in this paper has been implemented as part of
\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
tools.
Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
at \url{http://potassco.sourceforge.net}}.
The IG and parameters inference can be performed using the command
\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
The (possibly partial) inferred parametrization will be returned on the standard output.
The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
parameter to the command.

Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
our method infers the IG and parametrization given in \pref{fig:runningBRN}.
Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
undefined (because of the lack of cooperativity between $b$ and $c$).
In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
parameters with 3 potential values could not be inferred.
If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
admissible parametrizations.

The current implementation can successfully handle large PH models of BRNs found in the literature
such as an ERBB receptor-regulated G1/S transition model from \cite{Sahin09} which contains 20
components and 15 cooperative sorts, and a T-cells receptor model from \cite{Klamt06} which contains 40
components and 14 cooperative sorts\footnote{Both models are available as examples distributed with \textsc{Pint}.}.
For each model, IG and parameters inferences are performed together in less than a second
on a standard desktop computer.
%\footnote{Using a Dell Inspiron 1720 laptop, with an Intel Core 2 Duo CPU T5550 ($2 \times 1.83\text{GHz}$)
%and 3.9 Gib memory, on an Ubuntu 11.10 64-bits OS}
After removing the cooperations from these models (leaving only raw actions), the inferences allow to
determine 40 parameters out of 195 for the 20 components model, and 77 out of 143 for the 40 components model.
As we thus have an order of magnitude of respectively $10^{31}$ and $10^{73}$ admissible parametrizations,
these models would therefore be more efficiently studied as PH than as BRNs.
We note that the complexity of the method is exponential in the number of regulators of one
component and linear in the number of components.
% We note that despite its smaller size in term of components, the ERBB transmission model takes more time to be computed because the biggest cooperative sorts contain more processes (up to 32 processes) than in the T-cells receptor model (up to 8 processes).

A PH model can be built based on information found in the literature about the local influences between components.
The precision of this knowledge will determine the precision of the modeled activations and inhibitions,
and some information is likely to help in the representation of cooperations.
\end{comment}