### 2014-09-03 main.tex
intro

\documentclass[preprint]{elsarticle}
% vi:spell spelllang=en

\usepackage{hyperref}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}  % Maths
\usepackage{amsfonts} % Maths
\usepackage{amssymb}  % Maths
\usepackage{stmaryrd} % Maths (crochets doubles)

\usepackage{url}     % Mise en forme + liens pour URLs
\usepackage{array}   % Tableaux évolués

\usepackage{comment}

\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{pro}{Property~\ref{#1}}
\newrefformat{pps}{Proposition~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{sssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\newrefformat{ex}{Example~\ref{#1}}
\newrefformat{tb}{Table~\ref{#1}}
\def\pref{\prettyref}

\newdefinition{definition}{Definition}
\newdefinition{property}{Property}
\newdefinition{proposition}{Proposition}
\newdefinition{example}{Example}

\usepackage{tikz}
\newdimen\pgfex
\newdimen\pgfem
\usetikzlibrary{arrows,shapes,shadows,scopes}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}

\input{macros}
\input{macros-ph}
\input{macros-grn}

% Commandes À FAIRE
\usepackage{color} % Couleurs du texte
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
%\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}

\def\modMF#1{\textcolor{teal}{#1}}
\def\modLP#1{\textcolor{magenta}{#1}}
\def\modMM#1{\textcolor{blue}{#1}}
\def\modOR#1{\textcolor{olive}{#1}}

% Id est
\newcommand{\ie}{i.e.~}

% Césures
\hyphenation{pa-ra-me-tri-za-tion}
\hyphenation{pa-ra-me-tri-za-tions}

\begin{document}

\begin{frontmatter}

\title{\modLP{Identification of} Biological Regulatory Networks from Process Hitting models}
%\\\rewriteil{Or possibly: “Complementarity between PH, IG and Thomas modeling”}

\author[irccyn,nii]{Maxime Folschette}
\ead{Maxime.Folschette@irccyn.ec-nantes.fr}
\author[lri]{Lo\"ic Paulev\'e}
\author[nii]{Katsumi Inoue}
\author[irccyn,nii]{Morgan Magnin}
\author[irccyn]{Olivier Roux}

\address[irccyn]{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e - B.P. 92101 - 44321 Nantes Cedex 3, France.}
\address[nii]{National Institute of Informatics,\\
2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
\address[lri]{CNRS, Laboratoire de Recherche en Informatique UMR CNRS 8623\\
Universit\'e Paris-Sud, 91405 Orsay Cedex, France}

\begin{abstract}
\modMF{Maxime}
\modLP{Lo\"ic}
\modMM{Morgan}
\modOR{Olivier}

%\parindent 0.5cm
The Process Hitting (PH) \modLP{is a particular class of Indeterministic Asynchronous Automata
Network (or safe Petri nets) on which have been recently designed several static analyses for
dynamical properties that are tractable on very large systems.}

\modLP{In this paper, we address the formal identification of qualitative
models of Biological Regulatory Networks (BRNs) from PH models.}
First, the inference of the Interaction Graph (IG) from a PH model
\modLP{summarizes the signed influences between the components} that are effective for the dynamics.
Second, we provide the inference of all Ren\'e Thomas models of BRNs 
that are compatible with a given PH.
As the PH allows the specification of indeterministic interactions between
components, our inference emphasizes the ability of PH to deal with large BRNs with incomplete knowledge
on interactions, where Thomas' approach fails because of the combinatorics of parameters.

The inference of corresponding Thomas models is provided using Answer Set Programming,
which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.

%\todo{Rather name Kaufman as the creator of BRNs.}
%On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
%formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
%In addition to be able to use the existing results on this formalism,
%obtaining the Thomas representation of a PH model allows to iteratively refine a model.
%Linking this representation to PH would not only allow to use the existing results on this formalism,
%but also to iteratively refine the model of a system.

%This paper establishes formal links between these two formalisms, permitting to appreciate their
%complementary features.
\end{abstract}

\begin{keyword}
qualitative modeling \sep
model abstraction \sep
interaction graph \sep
René Thomas' parameters \sep
answer set programming
\end{keyword}

\end{frontmatter}

\input{parts-intro}

\section{Frameworks}\label{sec:frameworks}
\input{parts-ph}
\input{parts-grn}

\input{parts-infer-ig}
\input{parts-infer-param}
\input{parts-asp}
\input{parts-example}
\input{parts-discussion}

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

%\input{parts/contributions}

\end{document}